#Trends

Version: May 17 2023
Author: Danielle

##Setup

The setup files will get you ready for your analysis. It loads you libraries, needed tables, created directories for your outputs, and assigns some parameter values. 

Also assign max year of the analysis. Generally, the data in NatureCounts are 2 years behind. 

```{r setup}

max.year <- 2021 #needs changes with each analysis
ID<- "dethier" #chnage this to your user ID. You will be promted for a password. 

```

## Choose your CMMN station {#Setup3.1}

The number assigned to `t` will correspond to the row in the analysis parameters file ('anal.param') you want to analyze (each site has a separate row).

View the file to identify the row you want to select
```{r view}

## Import site-specific analysis parameters 
##This will need checked and updated before each analysis
anal.param <- read.csv("Data/CMMN_Analysis_ParameterValues.csv") 
View(anal.param)

```

Select your station. 
```{r chooseSite}

t <- 2

```

## Data manipulations {#Setup3.2}

Now that you have selected the CMMN site, we are going to clean the data. The cleaning scripts do the following: 

  - Load the data from the naturecounts database. You will be prompted for your password.
  - Subset data to min and max year based on analysis parameters file.
  - Calculate season-specific station windows (when a station normally operates).
  - Make list of sampling dates for zero-filling (based on *DET*)
  - Assign Species Code, and drop species that won't be analyzed (i.e., updates to taxonomies).
  - Use the superfile to filter to species and season of interest.
  - Assign UNEM For LPBO data only.
  - Save the manipulated data in the `Data` folder
  - Generates a species list for the analysis
  - create output tables for the analysis
  
First, assign 'u' your NatureCounts username and run the `prepare` scripts.

```{r data manip}
  
ID<-"dethier" 
source("01_prepare.R")
source("02_outputs.R")

```

## Data Analysis

From this point forward, we loop through species and season:

1) Zero-fill the data using the `events` matrix 

2) Apply species-specific filters to date (seasonal migration windows), abundance (mean 10 individuals/year), occurrence (mean 5 observation days/year), and drop years with no observations.

3) Run the trend analysis by looping through year periods (10-year periods)


```{r Change Defaults}

source("03_analysis.R")

```

NOTE: this step might not be necessary if you can input the obs variable right into the analysis. Leaving for now.

```{r pickObservationCountVar}


yr.data <- in.data %>%
  filter(ObservationCount > 0) %>%
  group_by(SpeciesCode, YearCollected, season) %>%
  summarize(totCount = sum(ObservationCount),
            nobs = n())



```
