---
output: html_document
editor_options: 
  chunk_output_type: console
---
output: html_document
editor_options: 
  chunk_output_type: console
---

#Effort

These scripts allow you to create the data outputs needed to inspect effort at each station. Since the data processing is different we will create a separate workflow. This file does not need updated every year, but perhaps every 5-10 years to determine if effort is changing systematically. 


##Setup

The setup files will get you ready for your analysis. It loads you libraries, needed tables, created directories for your outputs, and assigns some parameter values. 

Also assign max year of the analysis. Generally, the data in NatureCounts are 2 years behind. 

```{r setup}

max.year <- 2021 #needs changes with each analysis
min.year <- 2010 #how far back do you want the data to go? For most (all?) stations, the effort data started being collected in circa 2012. We will set the min date to 2010 so that we don't miss anything.  

```

## Choose your CMMN station {#Setup3.1}

The number assigned to `t` will correspond to the row in the analysis parameters file ('anal.param') you want to analyze (each site has a separate row).

View the file to identify the row you want to select
```{r view}

#This will need checked and updated before each analysis
anal.param <- read.csv("Data/CMMN_Analysis_ParameterValues.csv") 
View(anal.param)

```

##Select your station, or all stations in the analysis file if you wish to run a loop. 

**We may also choose to create a loop that runs through each stations and processes the data for the superfile analysis**
```{r chooseSite}

#Select a single site based on row number in anal.param
t <- 2 #test BBO

```

## Data manipulations {#Setup3.2}

Now that you have selected the CMMN site, we are going to clean the data. The cleaning scripts do the following: 

  - Load the data from the naturecounts database. You will be prompted for your password.
  - Assigns seasons (spring and fall)
  - Calculate season-specific station windows (when a station normally operates).
  - Make list of sampling dates for zero-filling.
  - Assign Species Code, and drop species that won't be analyzed (i.e., updates to taxonomies).
  - Assign UNEM For LPBO data only.
  - Save the manipulated data in the `Data` folder
  - Generates a species list for the analysis

This main differences from the analytically workflow `03-LoadManipSite` is that we are not using the superfile to filter to species and season of interest. We have also manually assigned the min and max year filters. 
  
First, assign 'u' your NatureCounts username and run the `prepare` scripts.

```{r data manip}
  
u<-"dethier"  #enter you NatureCounts ID
source("i_prepare.R")

```
