---
output: html_document
editor_options: 
  chunk_output_type: console
---
output: html_document
editor_options: 
  chunk_output_type: console
---

#Superfile

These scripts allow you to create the data outputs needed for the Superfile creation. Since the data processing is different we will create a septerate workflow. This file does not need updated every year, but perhaps every 5-10 years as migration timings shift due to climate change, foe example. 


# Set up Station Data {#Setup3}

```{r tidyr3, echo = FALSE, message = FALSE, warning = FALSE}

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy = FALSE)

```

##Setup

The setup files will get you ready for your analysis. It loads you libraries, needed tables, created directories for your outputs, and assigns some parameter values. 

Also assign max year of the analysis. Generally, the data in NatureCounts are 2 years behind. 

```{r setup}

max.year <- 2021 #needs changes with each analysis
min.year <- 2000 #how far back do you want the data to go. 
source("00_setup.R")

```

## Choose your CMMN station {#Setup3.1}

The number assigned to `t` will correspond to the row in the analysis parameters file ('anal.param') you want to analyze (each site has a separate row).

View the file to identify the row you want to select
```{r view}

view(anal.param)

```

##Select your station, or all stations in the analysis file if you wish to run a loop. 

**We may also choose to create a loop that runs through wach stations and processes the data for the superfile analysis**
```{r chooseSite}

t <- 10

```

## Data manipulations {#Setup3.2}

Now that you have selected the CMMN site, we are going to clean the data. The cleaning scripts do the following: 

  - Load the data from the naturecounts database. You will be prompted for your password.
  - Subset data to min and max year based on analysis parameters file.
  - Calculate season-specific station windows (when a station normally operates).
  - Make list of sampling dates for zero-filling.
  - Assign Species Code, and drop species that won't be analyzed (i.e., updates to taxonomies).
  - Assign UNEM For LPBO data only.
  - Save the manipulated data in the `Data` folder
  - Generates a species list for the analysis

This main differences from the analytically workflow `03-LoadManipSite` is that we are not using the superfile to filter to species and season of interest. We have also manually assigned the min and max year filters. 
  
First, assign 'u' your NatureCounts username and run the `prepare` scripts.

```{r data manip}
  
u<-"dethier"  #enter you NatureCOunts ID
source("a_prepare.R")

```

## Data Summary

We are now ready to create our data summaries.  

  - Loop through each species in the `species.list` which was create in the last step.
  - Zero fill the data frame using the `event.data` 
  - calculate the mean count per doy for each response variable in each season, where
  
    ObservationCount = DET
    ObservationCount3 = census
    ObservationCount4 = banding
    ObservationCount7 = census + banding
    
The code also calculates the proportion of years a species was detected on a specific day and the total number of years a site was monitored. 

```{r Change Defaults}

source("b_analysis.R")

```


## Plot Data

Now that we have the output for a station, we can plot the data using an interactive tool 

```{r plot}

mig.data<-read.csv(paste(out.dir, site, ".SuperData.csv", sep = ""))



```
