#Effort

Version: May 17 2023
Author: Danielle

These scripts allow you to create the data outputs needed to inspect effort at each station. Since the data processing is different we will create a separate workflow. This file does not need updated every year, but perhaps every 5-10 years to determine if effort is changing systematically. 

**Note that the raw data need to be download previously for this effort script to work. Please run the "03-LoadManipTrend.Rmd" to access and save the raw data in th 'Data' folder. 

## Choose your CMMN station 

The number assigned to `t` will correspond to the row in the analysis parameters file ('anal.param') you want to analyze (each site has a separate row).

View the file to identify the row you want to select. 
```{r view}

# This will need changed each analysis
max.year<-2021 

#enter your NatureCounts ID
ID="dethier" 

#This will need checked and updated before each analysis
anal.param <- read.csv("Data/CMMN_Analysis_ParameterValues.csv") 
View(anal.param)

```


```{r chooseSite}

#Select a single site based on row number in anal.param
t <- 3

```

## Data manipulations and plots 

This script was written to summarize and plot the effort data over time. The output file will be for all EffortMeasurements but the plots will be for just the EffortMeasurements that are standadized. 

```{r data manip}
  
source("effort.R")

```

These script will output a summary of effort variable in the 'out.dir' and plots in the 'plot.dir'. 
