#Superfile 

Version: May 17 2023
Author: Danielle

These scripts allow you to create the data outputs needed for the Superfile creation which are used for species classification. Since the data processing is different we will create a separate workflow. This file does not need updated every year, but perhaps every 5-10 years as migration timings shift due to climate change, for example. 

##Setup

The setup files will get you ready for your analysis. It loads you libraries, needed tables, created directories for your outputs, and assigns some parameter values. 

Also assign min and max year of the analysis. Generally, the data in NatureCounts are 2 years behind. 

```{r setup}

max.year <- 2019 #needs changes with each analysis
min.year.test <- 2000 #how far back do you want the data to go. The code will check that there are data this far back and correct if there is not. 
ID<-"dethier"

```

## Choose your CMMN station 

The number assigned to `t` will correspond to the row in the analysis parameters file ('anal.param') you want to analyze (each site has a separate row).

View the file to identify the row you want to select
```{r view}

#This will need checked and updated before each analysis
anal.param <- read.csv("Data/CMMN_Analysis_ParameterValues.csv") 
View(anal.param)

```

```{r chooseSite}

#Select a single site based on row number in anal.param
t <- 7 #test

```

## Data manipulations {#Setup3.2}

Now that you have selected the CMMN site, we are going to clean the data. The cleaning scripts do the following: 

This main differences from the analytically workflow `03-LoadManipTrend` is that we are not using the Superfile to filter to species and season of interest. We have also manually assigned the min and max year filters. 
  
We will import the `Raw_Data.csv` file from the database only if it hasn't been downloaded and saved in the `Data` folder.  

```{r data manip}

source("a_prepare.R")

```

## Data Summary

We are now ready to create our data summaries.  

  - Loop through each species in the `species.list` which was create in the last step.
  - Zero-fill the data frame using the `event.data` 
  - Drop any species from further analysis that falls below analysis thresholds for either criterion. The data are first filtered for the station widnow to eliminate high counts outside the regular count period. 
      Seasons where mean number of individuals/year is < 10 *based on DET*
      Seasons where mean number of observation days/year is < 5 *based on DET*
  - Calculate the mean count per doy for each response variable in each season, where
  
    ObservationCount = DET
    ObservationCount3 = census
    ObservationCount4 = banding
    ObservationCount7 = census + banding
    
  - Across all years, count up N of years for which each Day-of-year (doy) has a *DET* recorded.
  -	Calculate % of all years with data for that doy *based on DET*


```{r Change Defaults}

source("b_analysis.R")

```

## Classification Plots

    
-	Pick the doy at which % rises above 68%.  (Usually this is clear, but the odd day earlier may meet criterion, and vice versa, so some judgement needed.)

-	For each half-year, plot mean daily N vs. doy.  If possible, add a marker to graph showing start and end of coverage period. (Don’t have to be beautiful; just label with spp name and season. (Including the period before and after coverage period is helpful for assessing whether spp is present all summer or in winter.)

-	Examine plots to classify.  (Having done so much of this, I can do it quite quickly and with considerable confidence, so I’d be happy to take on this part if you could produce the graphs.)


```{r plot}

source("c_plot.R")

```

If you feeling ready (you feel the code is doing what you want it to do), you can use a loop and process all sites at once. 

```{r loop}

max.year <- 2019 
min.year.test <- 2000 
ID<-"dethier"
anal.param <- read.csv("Data/CMMN_Analysis_ParameterValues.csv") 

#remove sites that don't have data ready yet. 
anal.param<-anal.param %>% filter(site != "IBS") %>% filter(site != "OOT")

#assign length to the number of list in the list

for (j in 1:18) {
    
    t <- j
    
    source("a_prepare.R")
    source("b_analysis.R")
    source("c_plot.R")

}


```

